### 2016-07-06 ansi-isolation.tex
typos
Kia's suggestions and typos.

\subsection{Isolation Specifications}
\label{sec:ansi-isolation}

@@

We now describe specifications of standard isolation guarantees
expressed as constraints over trace well-formedness. For brevity and
convenience, we adopt few notations in this and subsequent sections.
An execution trace is destructed as ($\A$,$\visZ$) whenever individual
components of the pair are needed. Otherwise, it is written as $\E$.
Sometimes, the dot notation (\eg~$\E.\A$) is also used. Since $\A$ and
$\visZ$ are both sets, we lift the operations on sets to pairs of sets
when updating $\E$. For example, $\E' = \E \cup
(\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to $\E' = (\E.\A \cup
\{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.  When $\psi$ is a
formula, $\underE{\psi}$ denotes the interpretation of $\psi$ in the
context of the trace $\E$. Such interpretations are defined on a
case-by-case basis in Figs.~\ref{fig:rel-defs}
and~\ref{fig:ansi-isolation}.

@@

The ANSI SQL 92 standard requires \iso{Read Committed} isolation to
avoid the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. The {\sc rc}
specification (\C{RC})\footnote{In the following, we use small caps to abbreviate
  isolation levels (\emph{e.g.,} {\sc rc} for \emph{Read Committed}), and
  typewriter font for abbreviations of the specification of an isolation
  level (\emph{e.g.,} \C{RC} denotes the specification for {\sc rc} as
  given in Fig.~\ref{fig:ansi-isolation}).} is therefore a combination of these two
guarantees. The specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly
available replicated stores. On relational databases, however, {\sc
rc} has also come to be associated with the $\mathtt{MonotonicVis}$
guarantee.  Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$
are sufficient to reason about {\sc rc} isolation on relational stores
too. The combination of these guarantees with the {\sc sc} property of
relational stores (formalized in \S\ref{sec:store-consistency})
automatically leads to the monotonicity guarantee, which explains why
{\sc rc} comes with $\mathtt{MonotonicVis}$ on such stores regardless
of the implementation. On weakly consistent stores however,
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ do not imply
$\mathtt{MonotonicVis}$. A stronger isolation level called
\iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores.

>>>>

\subsection{Specifications}
\label{sec:ansi-isolation}

@@

We now describe specifications of standard isolation guarantees
expressed as constraints over trace well-formedness. For brevity and
convenience, we introduce some notations that are used in the following
sections.  An execution trace is destructed as ($\A$,$\visZ$) whenever
individual components of the pair are needed. Otherwise, it is written
as $\E$.  Sometimes, the dot notation (\eg~$\E.\A$) is also
used. Since $\A$ and $\visZ$ are both sets, we lift the operations on
sets to pairs of sets when updating $\E$. For example, $\E' = \E \cup
(\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to $\E' = (\E.\A \cup
\{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.  When $\psi$ is a
formula, $\underE{\psi}$ denotes the interpretation of $\psi$ in the
context of the trace $\E$. Such interpretations are defined on a
case-by-case basis in Figs.~\ref{fig:rel-defs}
and~\ref{fig:ansi-isolation}.

\subsubsection{Trace Relations}

@@

\subsubsection{Isolation Levels}

Fig.~\ref{fig:ansi-isolation} shows specifications of various
well-known isolation guarantees.  The ANSI SQL 92 standard requires
\iso{Read Committed} isolation to avoid the dirty reads phenomenon,
which is achieved by enforcing $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ guarantees. The {\sc rc} specification
(\C{RC})\footnote{In the following, we use small caps to abbreviate
isolation levels (\emph{e.g.,} {\sc rc} for \emph{Read Committed}),
and typewriter font for abbreviations of the specification of an
isolation level (\emph{e.g.,} \C{RC} denotes the specification for
{\sc rc} as given in Fig.~\ref{fig:ansi-isolation}).} is therefore a
combination of these two guarantees. The specification also agrees
with the description and implementation~\cite{bailishat,pldi15} of
{\sc rc} for highly available replicated stores. On relational
databases, however, {\sc rc} has also come to be associated with the
$\mathtt{MonotonicVis}$ guarantee.  Nonetheless, $\mathtt{AtomicVis}$
and $\mathtt{CommitVis}$ are sufficient to reason about {\sc rc}
isolation on relational stores too. The combination of these
guarantees with the {\sc sc} property of relational stores (formalized
in \S\ref{sec:store-consistency}) automatically leads to the
monotonicity guarantee, which explains why {\sc rc} comes with
$\mathtt{MonotonicVis}$ on such stores regardless of the
implementation. On weakly consistent stores however,
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ do not imply
$\mathtt{MonotonicVis}$. A stronger isolation level called
\iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores.

### 2016-07-06 reasoning.tex
Kia's suggestions and typos.

As mentioned earlier, our proof system is a rely-guarantee framework
that admits compositional reasoning by abstracting away environmental
interference as a rely relation. A conspicuous difference between a
standard development of rely-guarantee and ours is that, while the
former reasons in terms of program states (variable to value
bindings), we reason in terms of executions as captured by their
traces ($\E$). In particular, our rely ($R$) and guarantee ($G$)
relations relate executions (i.e., $R,\,Q\,\subseteq\,\E\times\E$),
and our pre- ($P$) and post- ($Q$) conditions are assertions over
executions (i.e., $P,\,Q\, : \E \rightarrow \Prop$). Our development
also facilitates state-based reasoning via the interpretation function
($\interp{\cdot}$) introduced in \S\ref{sec:syntax}, which interprets
the given set of effects as a state. For example, if a \emph{bi-state}
rely relation relates every pair of states $\sigma$ and $\sigma'$ such
that $\sigma'(X) \ge \sigma(X) \ge 0$, the corresponding
\emph{bi-execution} rely relation relates every pair of executions
$\E$ and $\E'$ such that $\interp{\E.\A}(X) \ge \interp{\E'.\A}(X) \ge
0$ . Assertions on states are also written similarly. For instance,
consider the post-condition $\C{B\,=\,k-a1-a2}$ of the program in
Fig.~\ref{fig:motiv-eg-1}. The corresponding assertion on the
post-state ($\lambda\sigma.~\sigma(\C{B}) \,=\, \C{k-a1-a2}$) asserts
that in all states resulting from executing the program, the value of
\C{B} is \C{k-a1-a2}. The equivalent execution-based assertion
($\lambda\E.~\interp{\E.\A}(\C{B})\,=\,\C{k-a1-a2}$) asserts that in
all executions of the program, the value of \C{B} is \C{k-a1-a2}.
However, having access to an execution facilitates assertions that go
beyond the state. An example is an invariant of \C{Wd1} described
abstractly in \S\ref{sec:motivation}, and reified as an
execution-based assertion below:

@@

The \rulelabel{RG-Conseq} rule define ways to strengthen or
weaken relations and assertions associated with the RG judgment of
transaction-bound expressions. Similar rules exist for
transaction-bound and top-level commands, but are not
shown.\footnote{The supplementary materials provide the complete set
  of rules.} As is the case with a standard rely-guarantee
formulation,the rules allow the pre-condition $P$ and the rely relation
$R$ to be strengthened, and the post-condition $Q$ (also, $C$ in the
case of expressions) and the guarantee relation $G$ to be weakened.
The most notable aspect of the \rulelabel{RG-Conseq} rules is that
they allow the trace invariant $\I$ to be strengthened. Considering
that $\I$ captures isolation properties, this means that a program
proven correct under weaker isolation levels is also correct under
stronger ones.  Parametricity over the trace invariant $\I$, combined
with the ability to strengthen $\I$ as needed, allows our proof system
to support a highly flexible proof strategy to prove programs correct
over various isolation variants. For example, programmers can
\emph{define} isolation guarantees of their choice (by defining $\I$
appropriately) and then prove programs correct assuming the guarantees
hold.  The soundness of strengthening $\I$ ensures that a program can
be safely executed on any system that offers isolation guarantees
at least as strong as those assumed.

>>>>

As mentioned earlier, our proof system is a rely-guarantee framework
that admits compositional reasoning by abstracting away environmental
interference as a rely relation. A conspicuous difference between a
standard development of rely-guarantee and ours is that, while the
former reasons in terms of program states (variable to value
bindings), we reason in terms of executions as captured by their
traces ($\E$). In particular, our rely ($R$) and guarantee ($G$)
relations relate executions (i.e., $R,\,G\,\subseteq\,\E\times\E$),
and our pre- ($P$) and post- ($Q$) conditions are assertions over
executions (i.e., $P,\,Q\, : \E \rightarrow \Prop$). Our development
also facilitates state-based reasoning via the interpretation function
($\interp{\cdot}$) introduced in \S\ref{sec:syntax}, which interprets
the given set of effects as a state. For example, if a \emph{bi-state}
rely relation relates every pair of states $\sigma$ and $\sigma'$ such
that $\sigma'(X) \ge \sigma(X) \ge 0$, the corresponding
\emph{bi-execution} rely relation relates every pair of executions
$\E$ and $\E'$ such that $\interp{\E.\A}(X) \ge \interp{\E'.\A}(X) \ge
0$ . Assertions on states are also written similarly. For instance,
consider the post-condition $\C{B\,=\,k-a1-a2}$ of the program in
Fig.~\ref{fig:motiv-eg-1}. The corresponding assertion on the
post-state ($\lambda\sigma.~\sigma(\C{B}) \,=\, \C{k-a1-a2}$) asserts
that in all states resulting from executing the program, the value of
\C{B} is \C{k-a1-a2}. The equivalent execution-based assertion
($\lambda\E.~\interp{\E.\A}(\C{B})\,=\,\C{k-a1-a2}$) asserts that in
all executions of the program, the value of \C{B} is \C{k-a1-a2}.
However, having access to an execution facilitates assertions that go
beyond the state. An example is an invariant of \C{Wd1} described
abstractly in \S\ref{sec:motivation}, and reified as an
execution-based assertion below:

@@

The \rulelabel{RG-Conseq} rule defines ways to strengthen or
weaken relations and assertions associated with the RG judgment of
transaction-bound expressions. Similar rules exist for
transaction-bound and top-level commands, but are not
shown.\footnote{The supplementary materials provide the complete set
  of rules.} As is the case with a standard rely-guarantee
formulation,the rules allow the pre-condition $P$ and the rely relation
$R$ to be strengthened, and the post-condition $Q$ (also, $C$ in the
case of expressions) and the guarantee relation $G$ to be weakened.
The most notable aspect of the \rulelabel{RG-Conseq} rules is that
they allow the trace invariant $\I$ to be strengthened. Considering
that $\I$ captures isolation properties, this means that a program
proven correct under weaker isolation levels is also correct under
stronger ones.  Parametricity over the trace invariant $\I$, combined
with the ability to strengthen $\I$ as needed, allows our proof system
to support a highly flexible proof strategy to prove programs correct
over various isolation variants. For example, programmers can
\emph{define} isolation guarantees of their choice (by defining $\I$
appropriately) and then prove programs correct assuming the guarantees
hold.  The soundness of strengthening $\I$ ensures that a program can
be safely executed on any system that offers isolation guarantees
at least as strong as those assumed.
